---
title: "100 R functions without installing extra packages"
author: "Thiyanga S. Talagala"
format:
  html:
    theme: journal
    toc: true
    toc_float: false
  pdf: default
  docx: default
project:
  type: website
  output-dir: docs
---

# Formatting

## `format`

To format numbers or strings according to a specific format

```{r}
format(12.3456789, digits = 2)
format(12.3456789, nsmall = 2)
format(1.5:5.5, trim = TRUE)
```

```{r}
df <- data.frame(id=c(1.1, 2.22, 3.333), letter=c("aaa", "bb", "c"))
df
format(df, justify = "left")
```

## `hexmode`

Print integers in hexmode

```{r}
values <- 10:100
as.hexmode(values)
```

# Common functions

## append

`x`: the vector the values are to be appended to

`values`: values to be included in the modified vectors

`after`: a subscript (index), after which the values are to be appended

```{r}
x1 <- seq(10, 50, by=10)
x2 <- 1:3
append(x1, values = x2, after = 3)
```

## 

# Matrix

```{r}
mat <- matrix(1:9, nrow =3)
```

## `det`

Calucate the determinant of a square matrix

```{r}
base::det(mat)
```

## `lower.tri/ upper.tri `

Check whether the elements belong to lower or upper triangle of a matrix

```{r}
base::lower.tri(mat)
```

```{r}
base::upper.tri(mat)
```

## `apply`

```{r}
apply(mat, 2, mean)
apply(mat, 2, max)
apply(mat, 1, max)
```

# Statistical functions

## `table`

Create a frequency table for a vector

```{r}
table(mtcars$cyl)
```

## IQR

```{r}
IQR(mtcars$mpg)
```

## `rank`

```{r}
x <- c(1, 1, 2, 3, 10, 1, 100)
rank(x)
```

# Graphics

## `boxplot`

```{r}
boxplot(mpg ~ cyl, data = mtcars, xlab = "Quantity of Cylinders",  
        ylab = "Miles Per Gallon", main = "Boxplot")

```


```{r}
boxplot(mpg ~ cyl, data = mtcars,   
        xlab = "Quantity of Cylinders",  
        ylab = "Miles Per Gallon",   
        main = "Boxplot Example",  
        notch = TRUE,   
        varwidth = TRUE,   
        col = c("green","yellow","red"),  
        names = c("High","Medium","Low")  
)  
```

## `curve`

```{r}
curve (5*x+4 , -2 , 2 , ylab='y')
curve (x^2 - 2 , add = TRUE , col = "red" , ylab = 'y ')
```

```{r}
curve (5*x^2 - 2, ylab = 'y ')
```

```{r}
expr <- cos
curve(expr ,-2*pi, 2*pi,  ylab='cos(x)', xlab='x')

```

## stem

```{r}
stem(mtcars$mpg)
```

# Other

## `sample`

x = a vector, from which to choose the sample

n = number of items to choose from 

replace = should sampling be with replacement?

size = size of the sample

prob = vector of probability weights for obtaining the elements of the vector being sampled

```{r}
data <- 1:10
sample(data)
sample(data, size = 5)
sample(data, size = 5, replace = TRUE)
sample(data, size = 5, replace = FALSE)
```